Redundant calibration as a complex optimization problem

### 2016-12-06 paper.tex
changed layout of paper

\subsection{Unpolarized Measurement Equation}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Eq.~\eqref{eq:vis_definition} is known as the unpolarized measurement equation. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are of course unknown and is what we try to estimate when we calibrate 
and image.

@@

@@

\subsection{Redundant Calibration}
\citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
that the Wirtinger framework can also be applied to redundant calibration.

@@

\subsection{Complex GN and LM}
In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{align}
\bmJ &= \frac{\partial \bv}{\partial \bz}, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{align}
In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator \citep{Sorber2012}.
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
where $\bD=\bI\odot\bJ^H\bJ$.

@@

\subsection{Analytic Expressions}
If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
\begin{equation}
\label{eq:Jacobian_red}
\bJ = \begin{bmatrix}
       \bM & \bN\\
       \conj{\bN} & \conj{\bM}
      \end{bmatrix},
\end{equation}
where
\begin{equation}
\bM =\begin{bmatrix}
      \bO & \bP
     \end{bmatrix},
\end{equation}
and 
\begin{equation}
\bN = \begin{bmatrix}
       \bQ & \bzero
      \end{bmatrix}.
\end{equation}
Moreover,
\begin{equation}
\left [ \bO  \right ]_{\alpha_{pq},j} = \begin{cases}
                                         y_{\phi_{pq}}\conj{g_q} & \textrm{if}~p=j\\
                                         0  & \textrm{otherwise} 
                                        \end{cases},
\end{equation}

@@

\subsection{Simplifying GN}

@@

\subsection{GN and LINCAL}
As we mentioned in the beginning of this section, the Wirtinger framework is not the only way of ensuring that $\bv$ is analytic in its argument. We could also 
have divided the problem in Eq.~\eqref{eq:least_squares_red} into its real and imaginary components after-which we solve for the real and imaginary parts of the parameters separately. There is a third way
of dealing with the issue. We can split the problem in Eq.~\eqref{eq:least_squares_complex} into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
This can be achieved by first rewriting Eq.~\eqref{eq:least_squares_complex} to
\begin{equation}
\label{eq:least_squares_lincal}
\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
\end{equation}
where $\bmath{\varrho}$ is the phase and amplitude parameter vector and is defined in Appendix~\ref{sec:lincal}. The Jacobian associated with 
Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. It should therefore be obvious to the reader that 
performing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in Appendix~\ref{sec:lincal} from first principles.
Therefore, a straightforward way of improving the convergence properties of stock standard LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.

\subsection{Diagonalization and StEFCal}
StEFCal (Statistically Efficient Calibration) is an alternating direction implicit method, in which the measurement equation (cite the equation earlier in paper) is linearized by assuming that the gains are known, but
that their conjugates are unknown (REMEMBER TO CITE ALL THE STEFCAL PAPERS MITCHELL ETC...). Under this assumption, the conjugates of the gains can be obtained trivially. Once the conjugates of the gains have been computed, the gains can be solved for. One can then alternate between computing the gains and their conjugates until convergence
is reached. By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative update step [MUST IMPROVE HERE]. \cite{Salvini2014} also show that
the convergence of the algorithm can be improved if for every even iteration the basic update step is modified as follows: the current gain estimate is replaced by the average of the current gain solution and the gain solution of the previous odd iteration.
StEFCal lowers the computational complexity of normal skymodel-based calibration from O$(P^3)$ to O$(P^2)$.

@@

>>>>

\subsection{Scalar Measurement Equation}
The observed visibility $d_{pq}$ measured by the baseline formed by antenna $p$ and $q$ is described by
\begin{equation}
\label{eq:vis_definition}
d_{pq} = g_{p}\conj{g_q}y_{pq} + n_{pq},
\end{equation}
where $g_{p}$ and $g_{q}$ are the direction independent gains associated with antenna $p$ and $q$, $y_{pq}$ denotes the true visibility that baseline $pq$ measured
and $n_{pq}$ is the thermal noise component. Eq.~\eqref{eq:vis_definition} is known as the unpolarized measurement equation. Conjugation is denoted by $\conj{(*)}$. The true value of $g_p$, $g_q$ and $y_{pq}$ are of course unknown and is what we try to estimate when we calibrate 
and image.

@@

NB:: ADD GN AND LM UPDATE RULES HERE

@@

\section{Redundant Calibration}

\subsection{Historical Overview}
Layered approach, LOGCAL, LINCAL, Stefan's conference paper, Noordam, other ASTRON paper, newest papers on redundant calibration, marthi paper etc...

\subsection{Real-Imag Framework}
Here we discuss the fact that LINCAL is just GN

OLD COPIED TEXT

As we mentioned in the beginning of this section, the Wirtinger framework is not the only way of ensuring that $\bv$ is analytic in its argument. We could also 
have divided the problem in Eq.~\eqref{eq:least_squares_red} into its real and imaginary components after-which we solve for the real and imaginary parts of the parameters separately. There is a third way
of dealing with the issue. We can split the problem in Eq.~\eqref{eq:least_squares_complex} into its phase and amplitude components and then solve for the amplitude and phase of the parameters separately.
This can be achieved by first rewriting Eq.~\eqref{eq:least_squares_complex} to
\begin{equation}
\label{eq:least_squares_lincal}
\min_{\bmath{\varrho}} \|\breve{\br}\| = \min_{\bmath{\varrho}} \|\breve{\bd} - \breve{\bv}(\bmath{\varrho})\|, 
\end{equation}
where $\bmath{\varrho}$ is the phase and amplitude parameter vector and is defined in Appendix~\ref{sec:lincal}. The Jacobian associated with 
Eq.~\eqref{eq:least_squares_lincal} is equal to Eq.~\eqref{eq:Jac_lin}. It should therefore be obvious to the reader that 
performing one iteration of LINCAL is equivalent to a basic GN update step; the GN update step was just re-derived in Appendix~\ref{sec:lincal} from first principles.
Therefore, a straightforward way of improving the convergence properties of stock standard LINCAL would be to simply add the LM damping factor to the basic LINCAL update step.

\subsection{Complex Framework}

\citet{Smirnov2015} recently proposed to use this Wirtinger framework instead of the classic real and imaginary approach to perform normal skymodel-based calibration. We will now show
that the Wirtinger framework can also be applied to redundant calibration.

@@

\subsubsection{Complex GN and LM}
In this section we discuss how we can use the GN and LM algorithms to solve Eq.~\eqref{eq:least_squares_complex} (i.e. estimate $\breve{\bz}$). The complex Jacobian is defined to be
\begin{equation}
\label{eq:Jacobian}
\bJ = \begin{bmatrix}
       \bmJ & \bmJ^*\\
       \conj{\bmJ}^* & \conj{\bmJ} 
      \end{bmatrix},
\end{equation}
where 
\begin{align}
\bmJ &= \frac{\partial \bv}{\partial \bz}, & \bmJ^* &= \frac{\partial \bv}{\partial \conj{\bz}}. 
\end{align}
In Eq.~\eqref{eq:Jacobian}, $\frac{\partial}{\partial \bz}$ and $\frac{\partial}{\partial \conj{\bz}}$ respectively denote the cogradient operator and the conjugate cogradient operator \citep{Sorber2012}.
The GN update step is now defined as:
\begin{equation}
\label{eq:GN_update}
 \Delta \breve{\bz} = (\bJ^H\bJ)^{-1}\bJ^H\breve{\br}.
\end{equation}
The LM update rule is very similar, the major difference being a damping factor $\lambda$ is introduced:
\begin{equation}
\label{eq:LM_update}
\Delta \breve{\bz} = (\bJ^H\bJ + \lambda\bD)^{-1}\bJ^H\breve{\br},
\end{equation}
where $\bD=\bI\odot\bJ^H\bJ$.

@@

\subsubsection{Analytic Expressions}
If we apply the definition in Eq.~\eqref{eq:Jacobian} to Eq.~\eqref{eq:least_squares_complex} we obtain the following analytic result:
\begin{equation}
\label{eq:Jacobian_red}
\bJ = \begin{bmatrix}
       \bM & \bN\\
       \conj{\bN} & \conj{\bM}
      \end{bmatrix},
\end{equation}
where
\begin{equation}
\bM =\begin{bmatrix}
      \bO & \bP
     \end{bmatrix},
\end{equation}
and 
\begin{equation}
\bN = \begin{bmatrix}
       \bQ & \bzero
      \end{bmatrix}.
\end{equation}
Moreover,
\begin{equation}
\left [ \bO  \right ]_{\alpha_{pq},j} = \begin{cases}
                                         y_{\phi_{pq}}\conj{g_q} & \textrm{if}~p=j\\
                                         0  & \textrm{otherwise} 
                                        \end{cases},
\end{equation}

@@

\subsubsection{Simplifying GN}

@@

\section{Results}

\subsection{Simulations}

\subsection{Alternating Direction Implicit Method}

%\subsection{Diagonalization and StEFCal}
StEFCal (Statistically Efficient Calibration) is an alternating direction implicit method, in which the measurement equation (cite the equation earlier in paper) is linearized by assuming that the gains are known, but
that their conjugates are unknown (REMEMBER TO CITE ALL THE STEFCAL PAPERS MITCHELL ETC...). Under this assumption, the conjugates of the gains can be obtained trivially. Once the conjugates of the gains have been computed, the gains can be solved for. One can then alternate between computing the gains and their conjugates until convergence
is reached. By correctly exploiting the symmetry inherent in the calibration problem this alternating approach can be replaced by a single iterative update step [MUST IMPROVE HERE]. \cite{Salvini2014} also show that
the convergence of the algorithm can be improved if for every even iteration the basic update step is modified as follows: the current gain estimate is replaced by the average of the current gain solution and the gain solution of the previous odd iteration.
StEFCal lowers the computational complexity of normal skymodel-based calibration from O$(P^3)$ to O$(P^2)$.

@@

NB:: NEED TO ADD SNR PLOTS HERE
PLUS IMAGE OF HESSIAN FOR THE DISCUSSION

\subsection{Conjugate Gradient Method}

\subsubsection{Spectral Condition Number}
kappa + itr plots
\subsubsection{Sparsity}
sparsity derivation
sparsity + computational complexity plots

\subsection{Comparison}
itr numbers plot and graph of $kP = P^{1.7}$.
